Learning Causally-Generated Stationary Time Series

### 2017-02-23 doc/paper.tex
Minor tweak.

Of particular interest for the current application are GP models that use non-parametric models of the power-spectra of the Gaussian Process (doubly non-parametric models) as they have the capacity to flexibly model signals with arbitrary spectral complexity. One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Bayesian nonparametric kernel learning, Wilson}, but it is not clear whether it is possible to build causality into such a construction. An alternative approach induces a non-parametric power spectrum by placing a Gaussian Process prior over the system's impulse response function  \citet{Tobar:2015:Learning_Stationary}. Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint (the impulse response in question should be thought of as an acausal spatial response). In this paper we revisit the GPCM and explicitly build in a causal constraint whilst retaining the flexibility of modelling the kernel nonparametrically. In addition to these modelling contributions, we develop enhanced variational inference and learning schemes for the causal and acausal variants of the GPCM -- collapsed variational bounds and structured approximating distributions -- that significantly improve statistical accuracy. These modelling and inferential contributions are demonstrated on synthetic and real world signals.

>>>>

Of particular interest for the current application are GP models that use non-parametric models of the power-spectra of the Gaussian Process (doubly non-parametric models) as they have the capacity to flexibly model signals with arbitrary spectral complexity. One approach employs a Dirichlet Process Mixture Model for the power spectra \cite{Bayesian nonparametric kernel learning, Wilson}, but it is not clear whether it is possible to build the causality constraint into such a construction. An alternative approach instead induces a non-parametric power spectrum by placing a Gaussian Process prior over the system's impulse response function  \citet{Tobar:2015:Learning_Stationary}. Critically, the so-called Gaussian Process Convolution Model (GPCM), does not incorporate a causality constraint (the impulse response in question should be thought of as an acausal spatial response). In this paper we revisit the GPCM and explicitly build in a causal constraint whilst retaining the flexibility of modelling the kernel nonparametrically. In addition to these modelling contributions, we develop enhanced variational inference and learning schemes for the causal and acausal variants of the GPCM -- collapsed variational bounds and structured approximating distributions -- that significantly improve statistical accuracy. These modelling and inferential contributions are demonstrated on synthetic and real world signals.
